---
title: "emileygz_OriginalHomeworkCode_02"
author: "Emiley Garcia-Zych"
date: "2023-09-20"
output:
  prettydoc::html_pretty:
  theme: cayman
  highlight: github
  toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# AN 588 Homework #2 "Titis In The Mist"

**Emiley Garcia-Zych**

"Every Saturday morning, at the same time, a primatologist goes and sits in the forest to listen for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number of calls she will hear in that time is exactly 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session."

**Variables**

one call is one event

time interval = 2 hours/1 sitting

the mean number of events per time interval, λ = 15

the number of calls per sitting = k

"The Poisson Distribution is often used to model open ended counts of independently occurring events, for example the number of cars that pass a traffic intersection over a given interval of time or the number of times a monkey scratches itself during a given observation interval. The *probability mass function* for the Poisson distribution is described by a single parameter, λ, where λ can be interpreted as the mean number of occurrences of the event in the given interval."
*C: I like that you provided an explanation of the Poisson distribution

## What is the probability that she will hear more than 8 calls during any given session?

```{r}
x <- 9 ## the first value over 8. 
l = 15 ## the mean number of events per time interval = 15.
cumulative <- 0 ## the combined probability
for(x in 9:31){ ## I used the range given later in the problem. The for loop will calculate the probability using the dpois function... dpois (x, lambda) = calculates Poisson distribution. 
probset <- dpois(x = x, lambda = l) ## store the probability of one case in the probset variable. 
cumulative <- cumulative + probset ## add probset variable to cumulative variable.
x <- x+1 ## increase # of calls being calculated by 1. 
}
print(cumulative)
```
*C: It is interesting how we both took this question differently and used different ways to calculate it. I used the ppois function applied to 8 and in fact got a very similar output with a small difference after the 3rd decimal, so I think both methods work. While I think your calculations may in fact be more exact to the question, if you wanted to make your code simpler you could in fact use ppois since the function gives a range of being more than the number you input, but that is entirely up to you in this case!
## What is the probability that she will hear no calls in a session?

```{r}
x <- 0 ## follow the same pattern as above. 
l = 15
probset <- dpois(x = x, lambda = l)
print(probset)
```
*C: I wonder if it may benefit the reader of your code if you explain the use of "print(probset)"? I see you use it throughout your code, and I think it could benefit your reader to just explain your thought process, otherwise your code is great!
## What is the probability that she will hear exactly 3 calls in a session?

k = 3 calls λ = 15 calls per sitting e = 2.718

```{r}
x <- 3 ## follow the same pattern as above. 
l = 15
probset <- dpois(x = x, lambda = l)
print(probset)
```

## Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.

```{r}
x <- 0:30
l = 15
probset <- dpois (x = x, lambda = l)
barplot(probset, names.arg = x, space = 0, col = "#6666FF", xlab = "number of calls per sitting", ylab = "probability", main = "Probability of Number of Calls Per Sitting") 
```
*C I really like how you colored your graph to make it more attractive and presentable!
## Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).

```{r}
simulation <- rpois (104, 15) ## use the rpois function to run a simulation. rpois (number of runs, lambda)
simulation ## print
```

## Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?

```{r}
hist(simulation, breaks = seq(0,30,1), probability = TRUE, col = "#6699FF", main = "Number of Calls Per Sitting (Simulation Values)", xlab = "number of calls", ylab = "probability") ## use the hist function to plot the data obtained in the part above. 
```
*C: I see you named your graph, and that inspired me to put a bit more into my own code to name my graph and even the axes! This looks really great!
The initial Poisson mass distribution is symmetrical and bell shaped. The simulation distribution is also symmetrical and bell shaped. Both graphs have a peak at around 15 calls.

*C: This code is really great and I can tell you applied a lot of what you learned in the modules to it. I hope any of my comments were helpful in improving anything, although overall I think you completed the homework assignment perfectly. Thank you fo rletting me take a look at your code and peer comment on it!