---
title: "emileygz_OriginalHomeworkCode_02"
author: "Emiley Garcia-Zych"
date: "2023-09-20"
output:
  prettydoc:: html_pretty
  theme: cayman
  highlight: github
  toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*HW Problem:* 
Every Saturday morning, at the same time, a primatologist goes and sits in the forest to listen for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number of calls she will hear in that time is exactly 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

*Variables*
one call is one event
time interval = 2 hours/1 sitting
the mean number of events per time interval, λ, is 15
the number of calls per sitting is k

"The Poisson Distribution is often used to model open ended counts of independently occurring events, for example the number of cars that pass a traffic intersection over a given interval of time or the number of times a monkey scratches itself during a given observation interval. The *probability mass function* for the Poisson distribution is described by a single parameter, λ, where λ can be interpreted as the mean number of occurrences of the event in the given interval."

## What is the probability that she will hear more than 8 calls during any given session?
```{r}
x <- 9
l = 15
cumulative <- 0
for(x in 9:31){
probset <- dpois(x = x, lambda = l)
cumulative <- cumulative + probset
x <- x+1
}
print(cumulative)
```

## What is the probability that she will hear no calls in a session?
```{r}
x <- 0
l = 15
probset <- dpois(x = x, lambda = l)
print(probset)
```
## What is the probability that she will hear exactly 3 calls in a session?

k = 3 calls
λ = 15 calls per sitting
e = 2.718

```{r}
x <- 3
l = 15
probset <- dpois(x = x, lambda = l)
print(probset)
```

## Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.

```{r}
x <- 0:30
l = 15
probset <- dpois (x = x, lambda = l)
barplot(probset, names.arg = x, space = 0, col = "orange", xlab = "number of calls per sitting", ylab = "relative frequency", main = "Probability Mass Function") 
```

## Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).

```{r}
simulation <- rpois (104, 15)
simulation
```

## Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?

```{r}
hist(simulation, breaks = c(0:30), probability = TRUE, col = "pink", main = "Number of Calls Per Sitting",
    xlab = "number of calls", ylab = "relative frequency")
```
